Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.

### 2016-10-02 paper_intersections.tex

\subsection{Intersection Line-Triangle (1d-2d)}
\label{sec:1d-2d}
Let us consider a line segment $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,\ t\in (0,1)
\end{equation}
and a triangle $S_2$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. 
\begin{lemma}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $\vc X$ on the $p$ and inside the triangle $S_2$. 
The barycentric coordinates of $\vc X$ on $S_2$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ has \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
The point $\vc X$ is inside of $S_2$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $\vc X$ on $S_2$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

The calculation of the intersection proceeds as follows:
\begin{enumerate}
    \item Compute or reuse \plucker coordinates and permuted inner products: $\pi_p$, $\pi_{s_i}$, $\pi_p \odot \pi_{s_i}$, for $i=1,2,3$.
    \item Compute barycentric coordinates $w_i,\ i=1,2,3$ using \eqref{eq:bary_centric}.
    \item If any $w_i$ is less than $\epsilon$, there is no intersection, return empty IP.  
    \item If all $w_i$ are greater than $\epsilon$, we set $\tau_{S_2} = (2, 0)$ for the IC.
    \item If one $w_i$ is less than $\epsilon$, intersection is on the edge $s_i$, we set $\tau_{S_2} =(1,i)$.
    \item If two $w_i$ are less than $\epsilon$, intersection is at the vertex $V_i$, we set $\tau_{S_2}=(0,i)$.
    \item \label{item:coplanar} If all $w_i$ are less than $\epsilon$, the line is coplanar with the triangle, both objects are 
        projected to the plane $x_i=0$ where $i$ is the index of the maximal component of the triangle's normal vector.
        Every pair $p$, $s_i$ is checked for an intersection on $S_2$ boundary either inside $s_i$ or at a vertex $V_i$, setting the topological info $\tau_T$ to 
        $(1, i)$ or $(0, i)$, respectively. At most two ICs are obtained.
    \item For each IC the barycentric coordinates $(1-t, t)$ on the line $p$ are computed according to \eqref{eq:line}. 
    \item If $t\in (-\epsilon, \epsilon)$ or $t\in (1-\epsilon, 1+\epsilon)$, we set $\tau_p = (0,0)$ or $\tau_p = (0,1)$, respectively.
    \item If $t\notin (-\epsilon, 1+\epsilon)$, the IC is eliminated.
\end{enumerate}

@@

\subsection{Intersection Line-Tetrahedron (1d-3d)}
\notePE{If we stick to $S_2$ in the section before, we can change line segment $p$ into $S_1$ here.}
In this section we consider an intersection of a line segment $p$ given by the parametric equation \eqref{eq:line_parametric} with a tetrahedron 
$S_3$. The used algorithm is based on the 1d-2d algorithm and closely follows \cite{platis_fast_2003}. Our modification takes into account
intersection with the line segment and consistently propagates topological position of ICs.

@@

  \DontPrintSemicolon
  \SetKw{and}{and}
  \SetKw{continue}{continue}
  \SetKw{break}{break}
  \KwIn{Tetrahedron $S_3$, line segment $p$.}
  \KwOut{List $I$ of ICs sorted along $p$.}  
  $I=\{\}$\; 
  \For{unmarked face $f$ of $S_3$}{
    $L$ = intersection($p,\ f$)\; \label{line:13-intersection}
    \lIf{ $L$ is none or degenerate}{\continue}
    \If{ $L$ is inside the edge $e$}{
        set $\tau_{S_3} = (1,e)$\;
        mark faces incident to $e$
    }
    \ElseIf{ $L$ is at the vertex $v$}{
        set $\tau_{S_3} = (0,v)$\;
        mark faces incident with $v$
    }
    append $L$ to $I$\; \label{line:13-append}
    \lIf{$\abs{I}=2$}{\break}
  }
  \lIf{$\abs{I}=1$ \and   $I$ is outside of $p$}{ erase $I$} \label{line:trimming}
  \ElseIf{$\abs{I}=2$}{    
    trim intersection with respect to the line segment $p$
  }
\end{algorithm}
Algorithm \ref{algo:13d} first computes line-face intersections for every face of $S_3$.
Tetrahedron has six edges, so 7 \plucker coordinates and 6 permuted inner products are computed at most. Precomputed coordinates and products are passed into the 1d-2d algorithm which is performed for the whole line $p$ (line \ref{line:13-intersection}).
If no IC is found, or coplanar case occurs in line-face computation, we continue to the next face. Note, ICs that would be created in coplanar case are to be found as ICs with the other faces, since they lie on edges.
Next, IC can be on an edge or at a vertex; then we set the correct topological position and mark the adjacent faces
to be skipped, since there cannot be another IC (and coplanar case has been checked already).
Finally at the line \ref{line:13-append}, we append the IC to the result and check whether the maximal amount of 
ICs has been reached.

After collecting line-tetrahedron ICs,
we do the line segment trimming from the line \ref{line:trimming} further. 
If we have only one IC, we check that it actually lies inside the line segment (otherwise, we throw it away).
If we have two ICs, and if both lie outside the line segment $p$, we eliminate both of them. 
If one of the ICs lies out of $p$,
we use the closest end point of the line segment instead and interpolate barycentric coordinates of the IC on $S_3$. The topological positions are updated as well.
The result of the algorithm is 0, 1, or 2 ICs, sorted by the parameter $t$ in the direction of the line $p$.

>>>>

\subsection{Intersection Line-Triangle (1d-2d)}
\label{sec:1d-2d}
Let us consider a line $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,
\end{equation}
on which a line segment $S_1$ is defined by $t\in [0,1]$ and a triangle $S_3$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. 
\begin{lemma}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $\vc X$ on the $p$ and inside the triangle $S_3$. 
The barycentric coordinates of $\vc X$ on $S_3$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ has \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
The point $\vc X$ is inside $S_3$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $\vc X$ on $S_3$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

The calculation of the intersection proceeds as follows:
\begin{enumerate}
    \item Compute or reuse \plucker coordinates and permuted inner products: $\pi_p$, $\pi_{s_i}$, $\pi_p \odot \pi_{s_i}$, for $i=1,2,3$.
    \item Compute barycentric coordinates $w_i,\ i=1,2,3$ using \eqref{eq:bary_centric}.
    \item If any $w_i$ is less than $\epsilon$, there is no intersection, return empty IP.  
    \item If all $w_i$ are greater than $\epsilon$, we set $\tau_{S_3} = (2, 0)$ for the IC.
    \item If one $w_i$ is less than $\epsilon$, intersection is on the edge $s_i$, we set $\tau_{S_3} =(1,i)$.
    \item If two $w_i$ are less than $\epsilon$, intersection is at the vertex $V_i$, we set $\tau_{S_3}=(0,i)$.
    \item \label{item:coplanar} If all $w_i$ are less than $\epsilon$, the line is coplanar with the triangle, both objects are 
        projected to the plane $x_i=0$ where $i$ is the index of the maximal component of the triangle's normal vector.
        Every pair $p$, $s_i$ is checked for an intersection on $S_3$ boundary either inside $s_i$ or at a vertex $V_i$, setting the topological info $\tau_{S_3}$ to 
        $(1, i)$ or $(0, i)$, respectively. At most two ICs are obtained.
    \item For each IC the barycentric coordinates $(1-t, t)$ on the line $p$ are computed according to \eqref{eq:line}. 
    \item If $t\in (-\epsilon, \epsilon)$ or $t\in (1-\epsilon, 1+\epsilon)$,
        we set the end point of $S_1$: $\tau_{S_1} = (0,0)$ or $\tau_{S_1} = (0,1)$, respectively.
    \item If $t\notin (-\epsilon, 1+\epsilon)$, the IC is eliminated.
\end{enumerate}

@@

\subsection{Intersection Line-Tetrahedron (1d-3d)}
In this section we consider an intersection of a line segment $S_1$, defined by an interval $t\in [0,1]$ of the line $p$ defined in \eqref{eq:line_parametric}, with a tetrahedron 
$S_3$. The used algorithm is based on the 1d-2d algorithm and closely follows \cite{platis_fast_2003}. Our modification takes into account
intersection with the line segment and consistently propagates topological position of ICs.

@@

  \DontPrintSemicolon
  \SetKw{and}{and}
  \SetKw{continue}{continue}
  \SetKw{break}{break}
  \KwIn{Line segment $S_1$ of line $p$, Tetrahedron $S_3$.}
  \KwOut{List $I$ of ICs sorted along $p$.}  
  $I=\{\}$\; 
  \For{unmarked face $f$ of $S_3$}{
    $L$ = intersection($p,\ f$)\; \label{line:13-intersection}
    \lIf{ $L$ is none or degenerate}{\continue}
    \If{ $L$ is inside the edge $e$}{
        set $\tau_{S_3} = (1,e)$\;
        mark faces incident to $e$
    }
    \ElseIf{ $L$ is at the vertex $v$}{
        set $\tau_{S_3} = (0,v)$\;
        mark faces incident with $v$
    }
    append $L$ to $I$\; \label{line:13-append}
    \lIf{$\abs{I}=2$}{\break}
  }
  \lIf{$\abs{I}=1$ \and   $I$ is outside of $S_1$}{ erase $I$} \label{line:trimming}
  \ElseIf{$\abs{I}=2$}{    
    trim intersection with respect to the line segment $S_1$
  }
\end{algorithm}
Algorithm \ref{algo:13d} first computes line-face intersections for every face of $S_3$.
Tetrahedron has six edges, so 7 \plucker coordinates and 6 permuted inner products are computed at most. Precomputed coordinates and products are passed into the 1d-2d algorithm which is performed for the whole line $p$ (line \ref{line:13-intersection}).
If no IC is found, or coplanar case occurs in line-face computation, we continue to the next face. Note, ICs that would be created in coplanar case are to be found as ICs with the other faces, since they lie on edges.
Next, IC can be on an edge or at a vertex; then we set the correct topological position and mark the adjacent faces
to be skipped, since there cannot be another IC (and coplanar case has been checked already).
Finally at the line \ref{line:13-append}, we append the IC to the result and check whether the maximal amount of 
ICs has been reached.

After collecting line-tetrahedron ICs,
we do the line segment trimming from the line \ref{line:trimming} further. 
If we have only one IC, we check that it actually lies inside $S_1$ (otherwise, we throw it away).
If we have two ICs, and if both lie outside $S_1$, we eliminate both of them. 
If one of the ICs lies out of $S_1$,
we use the closest end point of the line segment instead and interpolate barycentric coordinates of the IC on $S_3$. The topological positions $\tau_{S_1}$ and $\tau_{S_3}$ are updated as well.
The result of the algorithm is 0, 1, or 2 ICs, sorted by the parameter $t$ in the direction of the line $p$.
